\section{Introduction}
In this appendix is presented the new configuration file added to CPIM to support the configuration of the migration system.

\section{\textit{migration.xml}}
The template of the \textit{migration.xml} file is the following:

\begin{lstlisting}[language=XML, caption=migration.xml template]
<?xml version="1.0" encoding="UTF-8" standalone="no"?>
<migration>
    <zooKeeper>
        <type>...</type>
        <connection>...</connection>
        <range>...</range>
    </zooKeeper>
    <backup>
        <execute>...</execute>
        <type>...</type>
        <directory>...</directory>
        <prefix>...</prefix>
    </backup>
    <followCascades>...</followCascades>
</migration>
\end{lstlisting}

\newparagraph Three are the main section that can be configured:
\begin{enumerate}
\item \textbf{zooKeeper client}
\item \textbf{sequence number backup}
\item \textbf{follow cascades while build statements}
\end{enumerate}

\noindent The first two options are the most complex and are described in the following sections, the third option can assume the values of \textit{true} or \textit{false} but is \textit{optional} since is set to \textbf{false} by default.
In case the value is set to \textit{true}, the statement builders, when builds the statements from objects, will read the values for the \texttt{@CascadeType} annotation and, if necessary, builds the cascade statements and sends them to \textit{Hegira} in the correct execution order as described in \ref{sec:statements}.

\subsection{Configure the ZooKeeper client}
ZooKeeper client must be configured in the \textit{thread} of \textit{http} version as desbribed in \ref{sec:synch}.

\paragraph{Thread version} An example configuration would be the following one:
\begin{lstlisting}[language=XML, caption=ZooKeeper - thread type configuration]
<zooKeeper>
    <type>thread</type>
    <connection>localhost:2181</connection>
    <range>50</range>
</zooKeeper>
\end{lstlisting}

\noindent The \texttt{<connection>} tag is \textbf{required} and should contains an address in the form \texttt{host:port} which should be the host address and the port on which the ZooKeeper service is running.

\noindent The \texttt{<range>} is \textit{optional} since the default is \textbf{10} and is the default dimension of the range that \texttt{SeqNumberDispenser}(s) use to ask more sequence numbers to the synchronization system.

\paragraph{Http version} An example configuration would be the following one:
\begin{lstlisting}[language=XML, caption=ZooKeeper - http type configuration]
<zooKeeper>
    <type>http</type>
    <connection>http://server.com/hegira-api/zkService</connection>
    <range>50</range>
</zooKeeper>
\end{lstlisting}

\noindent In this case the \texttt{<connection>} tag should contains the API base path of the server in which the service is running.

\noindent For the \texttt{<range>} tag applies the same consideration made for the thread case.
 
\subsection{Configure a sequence number backup}
A sequence number backup can be configured either on a blob storage (when running on PaaS) or to file (when running to IaaS).

\newparagraph  The \texttt{<execute>} tag define if backup should or should not be performed, the possible values are \textit{yes} or \textit{no}. The tag is \textit{optional} since its default is \textbf{yes}.
It's supposed to be used for turning off the backup, in this case the configuration should be simply:
\begin{lstlisting}[language=XML, caption=Turning off sequence numbers backup]
<backup>
    <execute>no</execute>
</backup>
\end{lstlisting}

\newparagraph For the other configuration options, two cases must be distinguished:

\paragraph{Backup to Blob} An example configuration would be the following one:
\begin{lstlisting}[language=XML, caption=Backup to blob]
<backup>
    <type>blob</type>
    <prefix>SeqNumber_</prefix>
</backup>
\end{lstlisting}

\noindent The text in the \texttt{<prefix>} tag will be prefixed to each blob created to avoid file name conflicts.
This prefix is an \textit{optional} field since as default is set to \textbf{SeqNumber\textunderscore}.

\paragraph{Backup to file} An example configuration would be the following one:
\begin{lstlisting}[language=XML, caption=Backup to file]
<backup>
    <type>file</type>
    <directory>/backups</directory>
    <prefix>SeqNumber_</prefix>
</backup>
\end{lstlisting}

\noindent The \texttt{<directory>} tag is mandatory and must contain the path to the directory in which the backup file will be placed and retrieved in case they needs to be restored.

\noindent For the \texttt{<prefix>} tag, the same considerations made for backup to blob applies.

\section{Use CPIM without migration system}
The \textit{migration.xml} file is not necessary if the user won't use the migration system. If the file is not present inside the \texttt{META-INF} folder, the NoSQL service will interact only with the underlying persistence provider without instantiating any of the required classes.

\noindent This is possible due to the lazy initialization of those components that are initialized only the first time are actually used since are built with a singleton pattern.