In the last few years, the kind of data that applications needs to take care slightly changed from the data that applications used to handle in the years before, in particular the volume of the data to be both stored and served to the end-user is increased dramatically and moreover data has become unstructured.
Those changes in the data can be seen especially in Web application such as social networks in which handles a huge amount of user generated data.
In this scenario traditional DBMS solutions has been discussed since they not provide the performance that modern application require in handling such data. Hence, in recent years, NoSQL databases have emerged.

\noindent NoSQL databases try to address the new applications requirements in different ways, proposing different properties and characteristics. Each NoSQL database provides to its users a different API interface which, in general, does not abstract from the physical representation of the stored data and, thus, users need to deeply understand the internals of the NoSQL solution they are going to use. This is a problem that does not affect DBMS since SQL permit to manipulate data independently by the vendor implementation. 

\noindent The lack of a common language for NoSQL databases, require, when developing application, a clear understanding of the available NoSQL solutions, to be able to choose the right technology for the application requirements but, during the life cycle of the applications, many changes can arise but at this point, change the adopted NoSQL technology may become a problem. This problem is known as \textit{vendor lock-in}.  

\noindent To mitigate the problem this work proposes an extension of the CPIM library, which already try to address the vendor lock-in problem in PaaS environment, in order to provide to the user a way to abstract from the specific NoSQL technology used to store the data.
Many solutions has been proposed by both communities and industry in defining a common way to access different NoSQL technologies, we propose to use the one that seems to get most interest, especially by the industry, which is the JPA interface. We develop new extensions for Kundera, an ORM for NoSQL databases based on the JPA standard and we use it as abstraction layer in the NoSQL service of the CPIM library.
The complexity of using NoSQL databases is moved inside Kundera clients and this permit to the user to access several NoSQL technologies by using both a standard language, which is SQL, and a well known interface among Java developer, which is the JPA.

\noindent To achieve complete portability both of the code and of the data, this work also propose the integration of the migration and synchronization system \textit{Hegira}, in the CPIM. This permit to the user a way of moving its data from a NoSQL technology to another without experiencing application down time, and thus be able to change the NoSQL technology due, for example, to requirement changes.
 
\section*{Original Contributions}
This work include the following original contributions:
\begin{itemize}
\item two brand new Kundera clients, one for Google Datastore and one for Azure Tables;
\item the support, for the  NoSQL service of CPIM library, for the interaction with the migration system \textit{Hegira}, both for data synchronization and migration; 
\end{itemize}

\section*{Outline of the Thesis}
This thesis is organized as follows: 
\begin{itemize}
\item In Chapter \ref{chap:sota} is described the evolution of NoSQL. As a first introduction is discussed why in this years this technology have emerged over SQL solutions and what are the main differences among those technology, the second part aims to underline the lack of a common language for NoSQLs in contrast to DBMS in which SQL exists.
\item In Chapter \ref{chap:ps} we analyze the current available technologies for NoSQL databases and propose a work for extending the CPIM library to make its interaction with NoSQL world even easier and much more extensible. Furthermore we analyze the requirements of modern web application motivating the choice of integrating in the CPIM NoSQL service a mechanism of data migration.
\item Chapter \ref{chap:kundera} is dedicated to the develop of the two Kundera client extension that have been developed in order to support Google Datastore and Azure Tables that will be then used in CPIM as adapters for the relative database.
\item In Chapter \ref{chap:cpim} is presented the work made on CPIM. As a first step is described a modification in the CPIM NoSQL service aimed to integrate Kundera as unique persistence layer for NoSQL access using the standard JPA interface, the library was previously using several different JPA implementation one for each of the supported databases. Furthermore is discussed the extension of CPIM to include an interaction with the migration system \textit{Hegira}.
\item In Chapter \ref{chap:eval} are described the various tests that have been performed on the developed Kundera extensions, to guarantee the correctness of the operation and to provide a measurement of performance. Moreover is presented a test application developed to be able to test the migration of the data generated through the CPIM library through the NoSQL service.
\item In Chapter \ref{chap:conclusions} draws the conclusions on the entire work and proposes some possible future works.
\end{itemize}

