\section{Introduction}
In this appendix is descried the required procedure to build the benchmark project that contains the YCSB adapters, then are shown the required commands to be executed in order to execute the tho phases of an YCSB benchmark and the available values to be configured for each of the supported adapters.

\section{Preliminary operations}
In order to build the benchmark project, (available at \url{https://github.com/Arci/kundera-benchmark}) some libraries needs to be downloaded since are not available in any maven repository:
\begin{itemize}
\item Azure Table extension\\ \url{https://github.com/Arci/kundera-azure-table}
\item GAE Datastore extension \\\url{https://github.com/Arci/kundera-gae-datastore}
\end{itemize}

\noindent The Azure Table extension tests to run requires a reachable storage emulator on Windows so if this is not possible, skip tests by running \texttt{mvn clean install -DskipTests}.
 
\noindent Tests for the Datastore extension can be executed without any configuration as they are executed thought Google in-memory Datastore stub.

\noindent Also YCSB is not available in any maven repository, it must be downloaded (\url{https://github.com/brianfrankcooper/YCSB}) and installed locally, always through \texttt{mvn install}.

\newparagraph When all the required dependency for kundera-benchmark are resolved, is possible to install it with \texttt{mvn clean install} and then lunch the command \texttt{mvn dependency:copy-dependencies}, this will create a directory called \texttt{dependency} in the \texttt{target} directory containing all the jars of the dependencies. The \texttt{dependency} folder will be used for defining the classpath later on.

\section{Run tests for low-level API version}
\label{appendix:ycsb-low-level}
The two phases of the YCSB benchmark can be executed through the command:

\begin{lstlisting}[language=bash, caption=Run low-level API benchmarks, numbers=none]
java -cp KUNDERA-BENCHMARK-JAR-LOCATION:PATH-TO-DEPENDENCY-FOLDER/*
com.yahoo.ycsb.Client -t -db DATABASE-ADAPTER-CLASS-TO-USE
-P PATH-TO-WORKLOAD -P PATH-TO-PROPERTY-FILE 
-s -threads THREAD-TO-USE -PHASE > OUTPUT_FILE
\end{lstlisting}

\noindent where PHASE should be \texttt{load} for \textbf{load} phase or \texttt{t} for \textbf{transaction} phase.

\newparagraph  Available adapter classes are:
\begin{itemize}
\item \texttt{it.polimi.ycsb.database.AzureTableClient} for Azure Table;
\item \texttt{it.polimi.ycsb.database.DatastoreClient} for GAE Datastore;
\item \texttt{it.polimi.ycsb.database.KunderaHBaseClient} for Hbase.
\end{itemize}

\subsection{Property files}
As can be seen from the command, a property file must be specified. Properties files must provide information to locate the database to test when running the benchmarks on the low-level API versions.

\paragraph{Google Datastore} The available properties are:
\begin{itemize}
\item \texttt{url} (\textit{required});
\item \texttt{port} (\textit{optional}), default is 443;
\item \texttt{username} (\textit{required}), the username of an admin on the remote application;
\item \texttt{password} (\textit{required}), can be omitted if tests are against localhost.
\end{itemize}

\paragraph{Azure Table} The available properties are:
\begin{itemize}
\item \texttt{emulator} (\textit{optional}) [true\textbar false];
\item \texttt{account.name} (\textit{required}) if not using emulator, available from azure portal;
\item \texttt{account.key} (\textit{required}) if not using emulator, available from azure portal;
\item \texttt{protocol} (\textit{optional}) [http\textbar https], default is https.
\end{itemize}
\noindent If \texttt{emulator} is set to \textit{true} the remaining properties are ignored.

\paragraph{Hbase}
The properties must be configured inside the adapter class because, to be more accurate w.r.t. the Kundera client, connection cannot be done in the \texttt{init()} method. 

\noindent The properties can be set modifying the following constants:
\begin{itemize}
\item \texttt{node}, the master node location;
\item \texttt{port}, the master node port;
\item \texttt{zookeeper.node}, the node location for \texttt{hbase.zookeeper.quorum};
\item \texttt{zookeeper.port}, the node port for \texttt{hbase.zookeeper.property.clientPort}.
\end{itemize}

\noindent Since property file is not needed for Hbase, it does not need to be specified while running the benchmarks.

\section{Run tests for Kundera version}
\label{appendix:ycsb-kundera}
Two phases of the YCSB benchmark can be executed through the command:

\begin{lstlisting}[language=bash, caption=Run Kundera clients benchmarks, numbers=none]
java -cp KUNDERA-BENCHMARK-JAR-LOCATION:PATH-TO-DEPENDENCY-FOLDER/*
com.yahoo.ycsb.Client -t -db DATABASE-ADAPTER-CLASS-TO-USE
-P PATH-TO-WORKLOAD -s -threads THREAD-TO-USE -PHASE > OUTPUT_FILE
\end{lstlisting}

\noindent where PHASE should be \texttt{load} for \textbf{load} phase or \texttt{t} for \textbf{transaction} phase.

\newparagraph Available adapter classes are:
\begin{itemize}
\item \texttt{it.polimi.ycsb.database.KunderaAzureTableClient} for kundera-azure-table extension;
\item \texttt{it.polimi.ycsb.database.KunderaDatastoreClient} for kundera-gae-datastore extension;
 \item\texttt{it.polimi.ycsb.database.KunderaHBaseClient} for kundera-hbase extension.
\end{itemize}

\subsection{\textit{persistence.xml} configuration}
In the \textit{persistence.xml} file each persistence unit must be configured to locate the database to test.

\noindent The possible configurations are described in the appendix \ref{app:kconfig}.

\noindent Hbase configuration make use also of a datastore specific property file \texttt{hbase-properties.xml} in which can be configured the value for \texttt{hbase.zookeeper.quorum} and \texttt{hbase.zookeeper.property.clientPort}.